---
title: "R Notebook"
output: html_notebook
---
#Loading Libraries

```{r}
library(tidyverse)
library(janitor)
library(GGally)
library(modelr)
library(broom)
library(pROC)
library(glmulti)
library(caret)
```

# Reading in data

```{r}
orange_juice <- read_csv("data/orange_juice.csv") %>% 
  clean_names()
  
```

# Cleaning data

```{r}
orange_juice_clean <- orange_juice %>%
  mutate(purchase_mm = if_else(purchase == "MM", TRUE, FALSE),
         store_id = as.factor(store_id),
         special_ch = as.logical(special_ch),
         special_mm = as.logical(special_mm),
         pct_disc_mm = round(pct_disc_mm, digits = 2),
         pct_disc_ch = round(pct_disc_ch, digits = 2),
         disc_ch = if_else(disc_ch == 0, "no", "yes"),
         disc_mm = if_else(disc_mm == 0, "no", "yes")) %>% 
    mutate(weekof_purchase = sort(weekof_purchase, decreasing = FALSE)) %>% 
  mutate(quarter = ifelse(between(weekof_purchase, 227, 239), "q1", # taking weekof_purchase and dividing it into quarters of the year
                          ifelse(between(weekof_purchase, 240, 252), "q2",
                              ifelse(between(weekof_purchase, 253, 265), "q3", "q4")))) %>% 
  mutate(quarter = as.factor(quarter)) %>% 
  select(-store, -store7, -loyal_ch, -list_price_diff,-weekof_purchase, -purchase, -price_diff, -store_id)

  
```

#Checking for Alias Variables

```{r}
alias(purchase_mm ~ ., data = orange_juice_clean)
```

No alias variables present

# Splitting data into train and test

Choosing 80:20 split

```{r}
n_data <- nrow(orange_juice_clean)
test_index <- sample(1:n_data, size = n_data * 0.2)

test  <- slice(orange_juice_clean, test_index)
train <- slice(orange_juice_clean, -test_index)

#Check test and train data adds up
nrow(test) + nrow(train) == n_data
```

# Creating best logsistic regression model 

```{r}
glmulti_all_vars <- glmulti(
  purchase_mm ~ ., 
  data = train,
  level = 1,               
  method = "h",         
  crit = "bic",            
  confsetsize = 10,        
  plotty = F, 
  report = T,             
  fitfunction = "glm",     
  family = binomial(link = "logit")) 

summary(glmulti_all_vars)

```

purchase_mm ~ 1 + quarter + disc_mm + sale_price_mm + sale_price_ch is the best model

```{r}
glmulti_all_with_interaction <- glmulti(
  purchase_mm ~ 1 + quarter + disc_mm + sale_price_mm + sale_price_ch, 
  data = train,
  level = 2,               
  method = "h",            
  crit = "bic",            
  confsetsize = 10,        
  marginality = TRUE,      
  minsize = 6,             
  maxsize = 6,             
  plotty = F, 
  report = T,              
  fitfunction = "glm",    
  family = binomial(link = "logit")) 

summary(glmulti_all_with_interaction)
```

purchase_mm ~ 1 + quarter + disc_mm + sale_price_mm + sale_price_ch + sale_price_mm:disc_mm + sale_price_ch:sale_price_mm is the best model with interactions


# Identitfying best classifier

```{r}
orange_juice_model1_quarter <- glm(purchase_mm ~  quarter, data = orange_juice_clean, family = binomial(link = "logit"))

orange_juice_model2_disc_mm <- glm(purchase_mm ~  disc_mm, data = orange_juice_clean, family = binomial(link = "logit"))

orange_juice_model3_sale_price_mm <- glm(purchase_mm ~  sale_price_mm, data = orange_juice_clean, family = binomial(link = "logit"))

orange_juice_model4_sale_price_ch <- glm(purchase_mm ~  sale_price_ch, data = orange_juice_clean, family = binomial(link = "logit"))

orange_juice_model5_sale_price_mm_disc_mm <- glm(purchase_mm ~  sale_price_mm:disc_mm, data = orange_juice_clean, family = binomial(link = "logit"))

orange_juice_model6_sale_price_ch_sale_price_mm <- glm(purchase_mm ~  sale_price_ch:sale_price_mm, data = orange_juice_clean, family = binomial(link = "logit"))


```

## Adding predictions

```{r}
orange_juice_model1_quarter_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model1_quarter, type = "response")

orange_juice_model2_disc_mm_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model2_disc_mm, type = "response")

orange_juice_model3_sale_price_mm_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model3_sale_price_mm, type = "response")

orange_juice_model4_sale_price_ch_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model4_sale_price_ch, type = "response")

orange_juice_model5_sale_price_mm_disc_mm_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model5_sale_price_mm_disc_mm, type = "response")

orange_juice_model6_sale_price_ch_sale_price_mm_pred <- orange_juice_clean %>% 
  add_predictions(orange_juice_model6_sale_price_ch_sale_price_mm, type = "response")

```

## Creating ROC objects

```{r}
roc_obj_mod1 <- orange_juice_model1_quarter_pred %>%
  roc(response = purchase_mm, predictor = pred)

roc_obj_mod2 <- orange_juice_model2_disc_mm_pred %>%
  roc(response = purchase_mm, predictor = pred)

roc_obj_mod3 <- orange_juice_model3_sale_price_mm_pred %>%
  roc(response = purchase_mm, predictor = pred)

roc_obj_mod4 <- orange_juice_model4_sale_price_ch_pred %>%
  roc(response = purchase_mm, predictor = pred)

roc_obj_mod5 <- orange_juice_model5_sale_price_mm_disc_mm_pred %>%
  roc(response = purchase_mm, predictor = pred)

roc_obj_mod6 <- orange_juice_model6_sale_price_ch_sale_price_mm_pred %>%
  roc(response = purchase_mm, predictor = pred)

```

## Visualising ROC Curves

```{r}
roc_curve <- ggroc(
  data = list(
    mod1 = roc_obj_mod1, 
    mod2 = roc_obj_mod2,
    mod3 = roc_obj_mod3,
    mod4 = roc_obj_mod4,
    mod5 = roc_obj_mod3
  ), 
  legacy.axes = TRUE) +
  coord_fixed()

roc_curve
```

## Checking AUC Values

```{r}
auc(roc_obj_mod1)
auc(roc_obj_mod2)
auc(roc_obj_mod3)
auc(roc_obj_mod4)
auc(roc_obj_mod5)

```

Quarter is the best classifier with an area under the curve value of 0.6839

