---
title: "R Notebook"
output: html_notebook
---
#Loading in Libraries 
```{r}
library(rpart)
library(rpart.plot)
library(tidyverse)
library(GGally)
library(broom)
library(janitor)
library(modelr)
library(yardstick)

titanic_set <- read_csv('data/titanic_decision_tree_data.csv')

shuffle_index <- sample(1:nrow(titanic_set))

# shuffle the data so class order isn't in order - need this for training/testing split later on 
titanic_set <- titanic_set[shuffle_index, ]
```

```{r}
summary(titanic_set)
```

# Cleaning the titanic data

```{r}
clean_titanic_set <- titanic_set %>%
  filter(survived %in% c(0, 1)) %>%
  select(-c(X1, passenger_id, name, ticket, fare, cabin)) %>% 
  mutate_if(is.character, as.factor) %>% 
  mutate(pclass = factor(pclass, levels = c(3, 2, 1), labels = c("Lower", "Middle", "Upper")),
         survived = factor(survived, levels = c(0, 1), labels = c("did not survive", "survived")),
         age = as.numeric(as.integer(age)),
         age_status = as.factor(if_else(age <= 16, "child", "adult"))) %>% 
  na.omit()

clean_titanic_set
  
```

# Having an initial look at potentially siginificant variables

```{r}
ggpairs(clean_titanic_set, progress = FALSE)
```

"pclass", "sex" and "parch" could all potentially be important variables.  The above shows there is disparity between the box sizes in both pclass and sex with survived plots.

```{r}
dim(clean_titanic_set)
```

# Creating the test and training data - 20/80 split

```{r}
n_data <- nrow(clean_titanic_set)

#get an index
test_index <- sample(1:n_data, size = n_data * 0.2)

#get those rows from the actual data
titanic_test <- slice(clean_titanic_set, test_index)

# get the training dataset
titanic_train <- slice(clean_titanic_set, -test_index)
```

```{r}
titanic_test %>% 
  tabyl(survived)
```

```{r}
titanic_train %>% 
  tabyl(survived)
```

The 80/20 split was chosen as its one of the most common split percentages and there is enough data in the data set to do so. Additionally the 80/20 split follows Pareto principle (roughly 80% of the effects come from 20% of the causes for many events).

output tables show a fair balance


#Plotting decision tree
```{r}
titanic_fit <- rpart(survived ~ .,
                     data = titanic_train,
                     method = "class")
```



```{r}
rpart.plot(titanic_fit, yesno = 2)
```

The decision tree shows the important factors are sex, pclass, age and sib_sp in determining whether a passenger survives.

The first node shows that there is a 0.41 probability that a passenger did not survive from 100% of training data. The primary important factor is identified as sex. The second nodes in the hierarchy show that if you are not male you have a 0.76 probability of surviving (38% of training set) and if you are  
male you have a 0.21 probability of not surviving (62% of training set). 

Interesting points

- If you are female and pclass != lower then you have a 0.94 probablity of surviving (23% of the pop)
- if you are male and under age 13 and have 2 or more siblings you have a 0.94 probabilty of surviving (3% of the pop)
- if you are a female traveling lower class and below 28 years old with no siblings you have 0.71 chance of surviving (6% of pop)
- if you are male over 13 years old you have 0.17 probility of not surviving (57% of pop)
- if you are male, younger than 13 with 2 or more sibilings your probability of survival is the lowest (0.09) and makes up 2% of the population.


```{r}
titanic_test_pred <- titanic_test %>% 
  add_predictions(titanic_fit, type = "class")
```

```{r}
conf_matrix <- titanic_test_pred %>% 
  conf_mat(truth = survived, estimate = pred)
```

The confusion matrix shows that the decision tree identified correctly that 32 people survived and 85 did not.
It predicted incorrectly 20 people surviving and 5 not surviving.


