---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(lubridate)
library(tidyverse)
library(tsibble)
library(tsibbledata)
library(feasts)
library(nycflights13)

```

# 1. Dates and Times


```{r}
date1 <- ymd_hms("2017-04-22 22:01:18")
date2 <- ymd_hms("2001-11-07 10:09:56")
```

1. Extracting the different components (year, month, mday, yday, wday) from the variable date1.

```{r}
year(date1)
month(date1, label = TRUE, abbr = FALSE)
mday(date1)
yday(date1)
wday(date1)
```

2. - Adding 14 days to date1 and storing it in a variable called next_fortnight. 

- Taking away 10 years from date2 variable, and storing it in previous_decade.

```{r}
next_fornight <- date1 + days(14)
next_fornight

previous_decade <- date2 - years(10)
previous_decade
```

3. Creating a variable called this_year containing today’s date. Then creating an interval object which uses this_year and the previous_decade variable made above. Storing it in interval_years.

```{r}
this_year <- today()

interval_years <- previous_decade %--% this_year

interval_years
```

4. Changing the time zone of both your date variables to “America/New_York”, and call them nyc_date1 and nyc_date2.

```{r}
nyc_date1 <- with_tz(date1, tzone = "America/New_York")
nyc_date2 <- with_tz(date2, tzone = "America/New_York")

nyc_date1
nyc_date2
```

5. Using the flights data from the nycflights13 dataset and making a date-time column called departure_date from the year, month, day, hour, and minute variables in the flights tibble.

```{r}
flights %>% 
  mutate(departure_date = make_datetime(year, month, day, hour, minute))
```


# 2. Time series familiarity

6. 
```{r}
view(weather)

unique(weather$origin)
```

```{r}
weather_tibble <- weather %>% 
  select(temp, humid, precip, time_hour, origin)

```

The time index is preserved in the time_hour variable and a new observation is made every hour. Origin is the key as this together with time_hour, uniquely identifies the observation

7. Making the weather tibble into a weather tsibble

```{r}
weather_tsibble <- weather_tibble %>% as_tsibble(key = origin, index = time_hour)
```

8. 

```{r}
weather_month_year_cols <- weather_tsibble %>% 
  mutate(year = year(time_hour), month = month(time_hour, label = TRUE)) 

```

```{r}

weather <- weather_month_year_cols %>%
  index_by(date = ~ as_date(.)) 


ggplot(data = weather, aes(x = date, y = temp)) +
  geom_line(aes(colour = origin))
```



