---
title: "R Notebook"
output: html_notebook
---

# Loading in the required library 
```{r}

library(tidyverse)

```

# 1. Reading in the code_clan_tweets.csv file, returning the number of row and columns and listing variable names.

```{r}

code_clan_tweets <- read_csv("data/code_clan_tweets.csv")

view(code_clan_tweets) 

dim(code_clan_tweets) #Returns number of rows and columns

names(code_clan_tweets) #returns variable names


```

#2. A summary of how many likes each CodeClan tweet has.

```{r}

summarised_code_clan_tweets<- code_clan_tweets %>% 
  group_by(tweet_id) %>% # group the entries by tweet_id
  summarise(favorite_count)

boxplot(summarised_code_clan_tweets) 

```

# 3. Returning the number of CodeClan tweets that have been liked (>0 favourite count) minus tweets that were quoted.

```{r}
total_of_fave_tweets <- code_clan_tweets %>% 
  filter(!favorite_count == 0) %>% #reads as return tweets that don't equal zero only
  filter(!is_quote == TRUE) %>% # reads as return only tweets that weren't quoted
  summarise(total_fav_tweet_count = n()) # n() counts the number of tweets that were favourited

total_of_fave_tweets

```

# 4. Summary of the mean number of rewteets, split by platform type (source)

```{r}
mean_retweet_by_source <- code_clan_tweets %>% 
  filter(!is_quote == TRUE) %>% # reads as return only tweets that weren't quoted
  group_by(source) %>% # group the entries by source type
  summarise(mean_retweet_count = mean(retweet_count)) #summarises and creates new column that calculates mean of retweet_count

mean_retweet_by_source
```

# 5. The total number of favourite_count values, split by media_type in descending order.

```{r}

total_of_fave_tweets_by_media <- code_clan_tweets %>%
  select(media_type, favorite_count) %>% #isolates the media and favourite_count columns
  group_by(media_type) %>% #groups tweets by the media type used 
  summarise(total_tweet_count = n()) %>% #new column called total_tweet_count uses n() to count the no of favourited tweets 
  arrange(desc(total_tweet_count)) #total_tweet_count is arranged so the media_type with the highest total tweet count is top
  

  

total_of_fave_tweets_by_media
  

```


```{r}

#Also attempted to change all NA values to "text" in media_types 
# ifelse did not seem to work, says unused argument ("photo")

total_of_fave_tweets_by_media2 <- code_clan_tweets %>% 
 mutate(media_type_na = is.na(media_type)) %>%
  ifelse(media_type_na, "text", "photo") %>% 
  group_by(media_type_na) %>% 
  summarise(total_tweet_count = n()) %>%
  arrange(desc(total_tweet_count))

total_of_fave_tweets_by_media2
  
```

#6. Returning the mean number of characters CodeClan tweets contain

```{r}


mean_num_char <- code_clan_tweets %>%
  select(text) %>% # selected text column
  mutate(num_chars = nchar(text, allowNA = TRUE)) %>% # new column that counts no of chars (num_chars) of each tweet allowing NA values
  filter(is.na(num_chars) == FALSE) %>% # filters out the NA values
  summarise(mean_chars = mean(num_chars)) #creates new column that summarises and calculates the mean of the num_chars column 



mean_num_char



  
```


#7. Reading in the code_clan_info.csv file and joining it on to code_clan_tweets

```{r}

code_clan_info <- read_csv("data/code_clan_info.csv")


code_clan_joined <- inner_join(code_clan_tweets, code_clan_info, "tweet_id") #joined by tweet_id as this is a column in both datasets

view(code_clan_joined)

```


#8. Creating a new tibble called codeclan_hashtags - only the tweet_id and hashtags columns (in lowercase for analysis). NAs negated.

```{r}

codeclan_hashtags <- tibble(code_clan_joined$tweet_id, tolower(code_clan_joined$hashtags)) %>% 
  filter(!is.na(code_clan_joined$hashtags))

view(codeclan_hashtags)

#tibble() creates a tibble - both tweet_id and hashtags columns from code_clan_joined are used to create tibble. tolower() ensures all entries in hashtag are in lower case and then any hashtag entires that are NA values are filtered out using filter() and is.na().
```


#Extensions

```{r}

```






























