---
title: "R Notebook"
output: html_notebook
---

# Loading libraries 
```{r}
library(tidyverse)
library(modelr)
library(car)
library(GGally)
library(relaimpo)
library(janitor)
```

# Reading in data and exploring 

```{r}
avocado_data <- read.csv("data/avocado.csv") %>%
  clean_names()

glimpse(avocado_data)

view(avocado_data)

view(unique(avocado_data$region))
```

I have chosen to build an explanatory model for the average_price.

#Tidying avocado_data


```{r}
avocado_data_clean <- avocado_data %>% 
  dplyr::select(-c(x, date, year)) %>% 
  mutate(region = as.character(region)) %>% 
  mutate(region = if_else(str_detect(region,
 "Albany|BaltimoreWashington|BuffaloRochester|Boston|Charlotte|Detroit|HarrisburgScranton|HartfordSpringfield|Indianapolis|Jacksonville|Nashville|NewYork|Northeast|NorthernNewEngland|Orlando|Philadelphia|Pittsburgh|RichmondNorfolk|Roanoke|SouthCarolina|Southeast|Syracuse|Tampa|MiamiFtLauderdale|Atlanta"), "East", region)) %>% 
  mutate(region = if_else(str_detect(region,
 "Chicago|CincinnatiDayton|DallasFtWorth|Denver|Houston|Columbus|GreatLakes|Louisville|GrandRapids|Midsouth|NewOrleansMobile|Plains|StLouis"), "Midwest", region)) %>% 
   mutate(region = if_else(str_detect(region,
 "Boise|California|DallasFtWorth|Denver|Houston|LasVegas|LosAngeles|PhoenixTucson|Portland|RaleighGreensboro|Sacramento|SanDiego|SanFrancisco|Seattle|SouthCentral|Spokane|WestTexNewMexico"), "West", region)) %>% 
  mutate(region = as.factor(region))

glimpse(avocado_data_clean)
  
```

# Checking if TotalUS equates to all the other region categories added together
```{r}

avocado_data_clean %>% 
  filter(region == "TotalUS") %>%
  summarise(sum(total_volume))

avocado_data_clean %>% 
  filter(region != "TotalUS") %>%
  summarise(sum(total_volume))

# Doesn't appear to be the case 
```

```{r}
avocado_data %>% 
  ggplot(aes(x = region, y = average_price)) +
  geom_boxplot()

#Boxplot shows that region doesn't look like an important factor (not much separation between variable types) so it will be dropped 
```

```{r}
avocado_data_clean <- avocado_data_clean %>% 
  dplyr::select(-region)
```
```{r}
glimpse(avocado_data_clean)
```
```{r}
dim(avocado_data_clean)

avocado_trim <- avocado_data_clean %>% 
  drop_na()
dim(avocado_trim)
```

Looking at the distribution for average_price
```{r}
avocado_trim %>% 
  ggplot(aes(average_price))+
  geom_histogram()
```


# Checking for alias variables 

```{r}
alias(average_price ~ ., data = avocado_trim)
```

No alias variables in avocado_data_clean

# Finding correlation between variables and average_price

```{r}
ggpairs(avocado_trim)
```
Few of numeric variables show particularly strong correlation with average_price.

The highest correlation value is x4046 - 0.208 followed by total_volume - 0.193

The boxplot for type and average price shows separation between the two variable types.

#Building model

## 1st predictor

x4046 and average_price model 

```{r}
mod1a <- mod1a <- lm(average_price ~ x4046, data = avocado_trim)
summary(mod1a)
```

Residual standard error: 0.3939 on 18247 degrees of freedom
Multiple R-squared:  0.0434

```{r}
par(mfrow = c(2,2))
plot(mod1a)
```

The res v fit and scale-location plots show clear patterns, two distinct clusters in both.

total_volume and average_price model 

```{r}
mod1b <- mod1b <- lm(average_price ~ total_volume, data = avocado_trim)
summary(mod1b)
```

Residual standard error: 0.3951 on 18247 degrees of freedom
Multiple R-squared:  0.03715

Higher RSE and lower r-sq value than mod1a

```{r}
par(mfrow = c(2,2))
plot(mod1b)
```

Again the res v fit and scale-location plots show clear patterns, two disctinct clusters in both.

I will proceed with x4046

## 2nd Predictor

Adding residuals from mod1a and removing average_price and x4046

```{r}
avocado_resid <- avocado_trim  %>% 
  add_residuals(mod1a) %>% 
  dplyr::select(-c(average_price, x4046))
```

```{r}
ggpairs(avocado_resid)
```

There appears to be no other variable that correlates particularly with the residuals apart from type.

```{r}
mod2 <- lm(average_price ~ x4046 + type, data = avocado_trim)
summary(mod2)
```

Residual standard error: 0.316 on 18246 degrees of freedom
Multiple R-squared:  0.3843

RSE is down from mod1a and r-sq value is greater too indicating a better model

```{r}
par(mfrow = c(2,2))
plot(mod2)
```

Plots show points in vertical line due to categorical data type.

#Adding interaction

```{r}
mod3 <- lm(average_price ~ x4046 + type + x4046:type, data = avocado_trim)
summary(mod3)
```

Residual standard error: 0.3148 on 18245 degrees of freedom
Multiple R-squared:  0.3891

RSE is down from mod2 and r-sq value is greater too, indicating better model.

# Testing relative importance

```{r}
calc.relimp(mod3, type = "lmg", rela = TRUE)
```

type looks to be by far the most important factor in regards to average_price, lmg value is 92.5%
