---
title: "R Notebook"
output: html_notebook
---

```{r}
library(readr)
library(openxlsx)
library(readxl)
library(jsonlite)
library(CodeClanData)
```

# 1

```{r}
state_income_data <- read_csv("data/state_income_data.csv")

state_income_data

View(state_income_data)
```

# 2

# Wasn't sure what was exactly required for this question so I wrote in pseudo-code how to specify column names when using read_csv()

```{r}
read_csv("directory/file name.csv",
col_names = c("column_name1",
              "column_name2",
              "column_name3",
              "column_name4"))
```

# col_names() allows you to give the each column a name. I changed the 'vs' column to a logical by using cols(col_logical())

#3

```{r}

cars <- read_csv("data/cars.csv", cols(
  X1 = col_character(),
  mpg = col_double(),
  cyl = col_double(),
  disp = col_double(),
  hp = col_double(),
  vs = col_logical()
  ), col_names = TRUE
  )

cars

View(cars)
```


#4 scotland.txt was a txt file that used semi-colons to separate variables so read_delim was the most suitable to use. Specified 'delim = ";"' and 'skip = 2" (to make it easier to read)

```{r}

scotland <- read_delim("data/scotland.txt", delim = ";", skip = 2,
               col_names = c("Rank",
                             "Locality",
                             "Population",
                             "Status",
                             "Council Area"))


View(scotland)

```

#5 

# How to read text into a one element character vector in pseudo-code

```{r}

read_file("directory/file.txt")

```

#6 

# The read_csv function was most suitable to use, I needed to specify values for 'skip_empty_rows', 'n_max' and 'skip' for it to work 

```{r}

steam_200k_sample <- read_csv("data/steam-200k-sample.csv",
                                skip_empty_rows = TRUE, n_max = 10, skip = 4)

class(steam_200k_sample)

steam_200k_sample

View(steam_200k_sample)



```

#7 

# As the file is .tsv, read_tsv is the most suitable function

```{r}
historic_eu_payments <- read_tsv("data/historic_eu_payments.tsv")
View(historic_eu_payments)
```

#8

# How to read lines of text into a multi-element text vector in pseudo-code

```{r}
read_lines("directory/file.txt")
```


#9

# fromJSON allows the file to be read via URL, then parse_json() to interpret JSON file and convert to create make_me_ana1.  AFter this fromJSON can be used - which would return a data frame usually. In this case the 3 elements are all character vectors (same type) so it returns a list. 

```{r}
make_me_ana <- read_file("http://makemeanalyst.com/wp-content/uploads/2017/06/sample_json.txt")

#make_me_ana

make_me_ana1 <- parse_json(make_me_ana)

make_me_ana1

make_me_ana2 <- fromJSON(make_me_ana)

make_me_ana2



```








































































