---
title: "R Notebook"
output: html_notebook
---
# Loading libraries 
```{r}
library(tidyverse)
library(modelr)
library(car)
library(GGally)
library(relaimpo)
```
 
# Reading in data and exploring 

```{r}
houses <- read.csv("data/kc_house_data.csv")

glimpse(houses)

summary(houses)
```

```{r}
houses$waterfront
```


# Tidying houses data 

```{r}
houses_clean <- houses %>% 
  dplyr::select(-c("id", "date", "sqft_living15", "sqft_lot15", "zipcode")) %>% 
  mutate(waterfront = as.numeric(waterfront),
         condition = as.factor(condition),
         grade = as.factor(grade)) %>% 
  mutate(renovated = ifelse(yr_renovated > 0, TRUE, FALSE)) %>%
  dplyr::select(-yr_renovated)

glimpse(houses_clean)
```

The "id", "date", "sqft_living15", "sqft_lot15", "zipcode" variables are deemed not useful in creating a model for price and so are removed. 

"waterfront" is converted to numeric and "condition" and "grade" are converted to factors as they ordinal data types.

yr_renovated is mutated to a logical variable called renovated and yr_renovated is then removed.


# Checking for alias variables 

```{r}

alias(lm(price ~ ., data = houses_clean))

```

"sqft_basement" can be dropped as it is computed the same as "sqft_living" and computed also as not sqft_above - its an alias variable.


## Removing "sqft_basement" as it is alias variable
```{r}
houses_trim <- houses_clean %>% 
  dplyr::select(- sqft_basement)
```

## Rechecking for alias varibales 

```{r}

alias(lm(price ~ ., data = houses_trim))
```

No alias variables detected after removing "sqft_basement"

# Building a regression model

## Splitting the data into numeric and non-numeric so the data can 

```{r}
houses_trim_numeric <- houses_trim %>%
  select_if(is.numeric)

houses_trim_nonnumeric <- houses_trim %>%
  select_if(function(x) !is.numeric(x))

houses_trim_nonnumeric$price <- houses_trim$price

```

```{r}
houses_trim_numeric_plot <- ggpairs(houses_trim_numeric)
ggsave("houses_trim_numeric_plot.pdf")

```

```{r}
houses_trim_non_numeric_plot <- houses_trim_nonnumeric %>% 
  ggpairs()

houses_trim_non_numeric_plot
```

"sqft_living" and ""sqft_above" have strong correlations with price -  values of 0.702 and 0.606 respectively

"grade" boxplot in the non_numeric data shows some separation indicating potential for association as does "condition".


1st effect - "sqft_living"

```{r}
mod1a <- lm(price ~ sqft_living, data = houses_trim)
summary(mod1a)
```

Residual standard error: 261500 on 21611 degrees of freedom
Multiple R-squared:  0.4929

```{r}
par(mfrow = c(2,2))
plot(mod1a)
```

2nd effect 

```{r}
price_remaining_num_resid <- houses_trim_numeric  %>% 
  add_residuals(mod1a) %>% 
  dplyr::select(-c(price, sqft_living))
```

```{r}
price_remaining_num_resid_plot <- ggpairs(price_remaining_num_resid)
ggsave("price_remaining_num_resid_plot.pdf")
```

After subtracting off the systematic variation in price in the houses_trim_numeric, it appears that "lat" has the highest correlation value - 0.379

```{r}
mod2a <- lm(price ~ sqft_living + lat, data = houses_trim)
summary(mod2a)
```

Residual standard error: 241900 on 21610 degrees of freedom
Multiple R-squared:  0.566

Both figures are up from mod1a

```{r}
par(mfrow = c(2,2))
plot(mod2a)
```

Testing "grade"

```{r}
mod2b <- lm(price ~ sqft_living + grade, data = houses_trim)
summary(mod2b)
```

Residual standard error: 237500 on 21600 degrees of freedom
Multiple R-squared:  0.5817


Both r2 figures for mod2 models are up from mod1a but mod2b has a higher value and lower RSE.

```{r}
par(mfrow = c(2,2))
plot(mod2b)
```

price ~ sqft_living + grade

## Comparing the two models using ANOVA

```{r}
anova(mod2a, mod2b)
```


3rd Effect 

```{r}
price_remaining_nonnum_resid <- houses_trim  %>% 
  add_residuals(mod2b) %>% 
  dplyr::select(-c(price, sqft_living, grade))
```

```{r}
price_remaining_nonnum_resid_plot <- ggpairs(price_remaining_nonnum_resid)
ggsave("price_remaining_nonnum_resid.pdf")
```

when looking at the other options and result from the anova above, the 3rd predictor should be "lat"


```{r}
mod3 <- lm(price ~ sqft_living + grade + lat, data = houses_trim)
summary(mod3)
```

Residual standard error: 219000 on 21599 degrees of freedom
Multiple R-squared:  0.6443

RSE down and r-squared up -  meaning model is better fit 

price ~ sqft_living + grade + lat

#4th effect

going back to the houses_trim_non_numeric_plot "condition" also looks to have separtion in the boxplot meaning possible association.


```{r}
mod4 <- lm(price ~ sqft_living + grade + lat + condition, data = houses_trim)
summary(mod4)
```

Residual standard error: 215100 on 21595 degrees of freedom
Multiple R-squared:  0.657

RSE down and r-squared up -  meaning model is better fit than mod3

Extension 

```{r}
calc.relimp(mod4, type = "lmg", rela = TRUE)

#grade is the most important factor
```

