---
title: "R Notebook"
output: html_notebook
---

# Loading in libraries 

```{r}
library(tidyverse)
library(forcats)
library(janitor)
library(scales)
library(fable)
library(fabletools)
library(tsibble)
```

# Reading in the data and then cleaning variable names

```{r}
geo_codes_and_labels <- read_csv("geography_codes_and_labels_hb2014_01042019.csv") %>% 
  clean_names()

cancer_incidence <- read_csv("opendata_inc9418_hb.csv") %>% 
  clean_names()

```

# Exploring the data

```{r}

head(geo_codes_and_labels)
view(geo_codes_and_labels)

head(cancer_incidence)
summary(cancer_incidence)
```

## The data appears to be in a wider form than preferred for analysis but I've feel I can perform the analysis I need without formatting it too much.

```{r}
unique(cancer_incidence$hb)

view(borders_cancer_incidence)
```

# Recoding the NHS Board names to allow for easier analysis

```{r}
recoding_cancer_data <- cancer_incidence %>% 
  mutate(hb =case_when(
    hb == "S08000015" ~ "NHS Ayrshire and Arran",
    hb == "S08000016" ~ "NHS Borders",
    hb == "S08000017" ~ "NHS Dumfries and Galloway",
    hb == "S08000029" ~ "NHS Fife",
    hb == "S08000019" ~ "NHS Forth Valley",
    hb == "S08000020" ~ "NHS Grampian",
    hb == "S08000031" ~ "NHS Greater Glasgow and Clyde",
    hb == "S08000022" ~ "NHS Highland",
    hb == "S08000032" ~ "NHS Lanarkshire",
    hb == "S08000024" ~ "NHS Lothian",
    hb == "S08000025" ~ "NHS Orkney",
    hb == "S08000026" ~ "NHS Shetland",
    hb == "S08000030" ~ "NHS Tayside",
    hb == "S08000028" ~ "NHS Western Isles",
    TRUE ~ hb))

recoding_cancer_data

```

According to the NHS website, as their is a strong link between age and risk of cancer, a high crude rate is a good indication of a high proportion of elderly people within a population. It is therefore acceptable to say crude rate values will be valuable for NHS Borders to look at when planning for provision of cancer treatment services.

# Average Crude Rates for all cancer types over 24 years period, comparison by NHS Board 

```{r}
crude_rate_comparison <- recoding_cancer_data %>%
  filter(sex == "All",
         cancer_site == "All cancer types") %>% 
    group_by(hb) %>%
    summarise(aver_crude_rate = mean(crude_rate)) %>%
  mutate(hb = fct_reorder(hb, aver_crude_rate)) %>% 
  ggplot() +
  coord_flip() +
  aes(x = hb , y = aver_crude_rate) +
  geom_bar(stat = "identity") +
  labs( y = "Average Crude Rate",
       title = "The Average Crude Rate for all Cancer Types by NHS Board",
       subtitle = "1994 - 2018")
  
  

crude_rate_comparison 
  
```

The Borders reports the 2nd highest average crude rate over the 24 year period.


# How crude rate has varied over time for all cancer types within NHS Borders catchment area

```{r}
cr_borders_over_time <- recoding_cancer_data %>%
  filter(sex == "All",
         cancer_site == "All cancer types", hb == 'NHS Borders') %>% 
ggplot() +
  aes(x = year, y = crude_rate) +
  geom_line() +
  geom_point()+
  scale_x_continuous(breaks = breaks_extended(14))+
  labs(x = "Year",
       y = "Crude Rate",
       title = "Crude Rate for All Cancer types in NHS Borders",
       subtitle = "1994 - 2018")

cr_borders_over_time
```

Their appears to be an upwards trend of crude rate over the 24 year period.


# All recorded incidences of cancer reported in Borders by year

```{r}
borders_cancer_incidence <- cancer_incidence %>%
  filter(hb == "S08000016")
  
```

# The Number of Incidences of Cancer (all types) in NHS Borders

```{r}
all_cancer_types <- borders_cancer_incidence %>% 
  filter(cancer_site == "All cancer types",
         sex == "All") %>% 
  group_by(year) %>% 
  summarise(incidences_all_ages) %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages) +
  geom_line() +
  geom_point()+
  scale_x_continuous(breaks = breaks_extended(14))+
  labs(x = "Year",
       y = "Incidences",
       title = "The Number of Incidences of Cancer (all types) in NHS Borders",
       subtitle = "1994 - 2018")

all_cancer_types
```
  
# Splitting the above by sex 
  
```{r}
all_cancer_types_by_sex <- borders_cancer_incidence %>% 
  filter(cancer_site == "All cancer types",
         sex != "All") %>% 
  group_by(year, sex) %>% 
  summarise(incidences_all_ages) %>% 
  ggplot() +
  aes(x = year, y = incidences_all_ages, colour = sex) +
  geom_line() +
  geom_point()+
  scale_x_continuous(breaks = breaks_extended(10))+
  scale_y_continuous(breaks = breaks_extended(8)) +
  labs(x = "Year",
       y = "Incidences",
       title = "The Number of Incidences of Cancer (all types) split by Sex in NHS Borders",
       subtitle = "1994 - 2018")

all_cancer_types_by_sex
```

# The top 10 cancer types in NHS Borders 

```{r}
top_cancer_types <- borders_cancer_incidence %>% 
  filter(cancer_site != "All cancer types", sex == "All") %>% 
  group_by(cancer_site) %>%
  summarise(num_incidences = sum(incidences_all_ages)) %>%
  arrange(desc(num_incidences)) %>%
  mutate(cancer_site = fct_reorder(cancer_site, num_incidences)) %>% 
  head(10) %>% 
  ggplot() +
  aes(x = cancer_site , y = num_incidences) +
  geom_bar(stat = "identity") +
  coord_flip()+
  scale_y_continuous(breaks = breaks_extended(8)) +
  labs(x = "Cancer Type",
       y = "Incidences",
       title = "The Top Ten Cancer Types recorded in NHS Borders by Number of Incidences",
       subtitle = "1994 - 2018")
  

 
top_cancer_types
```

# The top 5 cancer types in NHS Borders (females)


```{r}
top_cancer_types_sex <- borders_cancer_incidence %>% 
  filter(cancer_site != "All cancer types", sex == "Female") %>% 
  group_by(cancer_site, sex) %>%
  summarise(num_incidences = sum(incidences_all_ages)) %>%
  arrange(desc(num_incidences)) %>%
  mutate(cancer_site = fct_reorder(cancer_site, num_incidences)) %>% 
  head(5) %>% 
  ggplot() +
  coord_flip() +
  aes(x = cancer_site , y = num_incidences) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(breaks = breaks_extended(8)) +
  labs(x = "Cancer Type",
       y = "Incidences",
       title = "The Top Ten Cancer Types recorded in NHS Borders by Number of Incidences",
       subtitle = "1994 - 2018")

top_cancer_types_sex 
```

# The top 5 cancer types in NHS Borders (males)

```{r}
top_cancer_types_sex <- borders_cancer_incidence %>% 
  filter(cancer_site != "All cancer types", sex == "Male") %>% 
  group_by(cancer_site, sex) %>%
  summarise(num_incidences = sum(incidences_all_ages)) %>%
  arrange(desc(num_incidences)) %>%
  mutate(cancer_site = fct_reorder(cancer_site, num_incidences)) %>% 
  head(5) %>% 
  ggplot() +
  coord_flip() +
  aes(x = cancer_site , y = num_incidences) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(breaks = breaks_extended(8)) +
  labs(x = "Cancer Type",
       y = "Incidences",
       title = "The Top Ten Cancer Types recorded in NHS Borders by Number of Incidences",
       subtitle = "1994 - 2018")

top_cancer_types_sex 
```

```{r}
view(borders_cancer_incidence)

unique(borders_cancer_incidence$cancer_site)
```



#Forecasting Crude Rate for Top 3 Cancer types in the Borders

```{r}
 

top3_cancer_types_cr <- borders_cancer_incidence %>% 
  filter(cancer_site == "Non-melanoma skin cancer" |
         cancer_site == "Breast" |
         cancer_site == "Basal cell carcinoma of the skin",
         sex == "All") %>% 
  group_by(year) %>%
  summarise(crude_rate = crude_rate, cancer_site) %>%
  arrange(desc(crude_rate)) %>% 
  mutate(year = as.numeric(year))

top3_cancer_types_cr

top3_cancer_types_cr_tsibble <- as_tsibble(top3_cancer_types_cr, index = year, key = cancer_site)
  
cancer_site_1 <- top3_cancer_types_cr_tsibble %>% 
  select(cancer_site, crude_rate, year)

cancer_site_1 %>% autoplot()

```

```{r}
fit <- cancer_site_1 %>%
  fabletools::model(
    arima = ARIMA(crude_rate),
    drift = RW(crude_rate ~ drift())
  )
fit
```


```{r}
forecast_5years <- fit %>%
  fabletools::forecast(h = "5 years")
forecast_5years 
```

```{r}
forecast_5years %>%
  autoplot(cancer_site_1, level = NULL) +
  ggtitle("Forecasts for Top 3 Cancer Types in NHS Borders Catchment") +
  xlab("Year") +
  guides(colour = guide_legend(title = "Forecast"))
```


The forecast models show a similar trend to the "Crude Rate for All Cancer types in NHS Borders" plot (fluctuating but ultimately upward trending) in regards to the top 3 cancer types.
