---
title: "R Notebook"
output: html_notebook
---
# Loading libraries

```{r}
library(tidyverse)
library(skimr)
library(janitor)
library(MASS)
library(ggpubr)
```

# Reading data and cleaning variable names

```{r}
mobile_details <- read_csv("data/20190928-items.csv")
clean_mobile_details <- clean_names(mobile_details)
```

# Viewing data

```{r}
view(clean_mobile_details)
```

# Brand with the highest number of phones in dataset

```{r}

most_phones_brand <- sort(table(clean_mobile_details$brand), decreasing = TRUE)[1]
most_phones_brand
```

# Plotting a distribution of phone ratings as a probability density and overlaying a fitted normal distribution for Samsung

Calculating mean and SD of Samsung phone ratings

```{r}
samsung_rating_stats <-clean_mobile_details %>% 
  filter(brand == "Samsung") %>% 
summarise(mean_rating = mean(rating),
          sd_rating = sd(rating),
          num = n())

samsung_rating_stats
```

Creating histogram and normal distribution plot

```{r}
samsung_dist_plot <- clean_mobile_details %>%
  filter(brand == "Samsung") %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white") +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = samsung_rating_stats$mean_rating, 
      sd = samsung_rating_stats$sd_rating
    ),
    col = "green"
  )

samsung_dist_plot
```

The plot shows that the distribution is skewed to the left so is not a good model of these ratings.

Using shapiro test to test normallity also confirms this as the p-value is below 0.05:

```{r}
samsung_sp <- clean_mobile_details %>% 
  filter(brand == "Samsung")
```

```{r}
shapiro.test(samsung_sp$rating)
```


#Calculating within one, two and three standard deviations of the mean 

Within one 

```{r}
samsung_sp %>%
  filter(rating >= samsung_rating_stats$mean_rating - samsung_rating_stats$sd_rating) %>%
  filter(rating <= samsung_rating_stats$mean_rating + samsung_rating_stats$sd_rating) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_sp))
```

Within two 

```{r}
samsung_sp %>%
  filter(rating >= samsung_rating_stats$mean_rating - (2 * samsung_rating_stats$sd_rating)) %>%
  filter(rating <= samsung_rating_stats$mean_rating + (2 * samsung_rating_stats$sd_rating)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung_sp))
```

Within three

```{r}
samsung_sp %>%
  filter(rating >= samsung_rating_stats$mean_rating - (3 * samsung_rating_stats$sd_rating)) %>%
  filter(rating <= samsung_rating_stats$mean_rating + (3 * samsung_rating_stats$sd_rating)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung_sp))
```

Normal distributions follow the 68-95-99.7 rule - the proportions of values lying within one-, two- and three standard deviations of the mean for samsung ratings are 0.7934509, 0.9193955, 0.9798489 (79%, 91% and 97.98%). This again demonstrates that normal distribution doesn't provide a good model.




