---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(infer)
library(janitor)
```

```{r}
ames <- read.csv("data/ames.csv") %>% 
  clean_names()
```

```{r}
glimpse(ames)
```

# Investigating the distribution of lot_area

```{r}
ames %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white")
```


The distribution is fairly normal in shape, slightly skewing right.


# Creating a sample:

```{r}
sample_1000 <- ames %>% 
  rep_sample_n(size = 1000, reps = 1)

sample_1000
```

# Checking sample

```{r}
sample_1000 %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white")
```

# Bootstrapping sample using infer and visualising

```{r}
infer_resample <- sample_1000 %>% 
  specify(response = lot_area) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")

infer_resample
```

```{r}
infer_resample_visualised <- infer_resample %>% 
  visualise()

infer_resample_visualised

```

# Using bootstrap distribution to calculate a 95% CI for mean(lot_area) and visualise it on the distribution

```{r}
infer_ci_95 <- infer_resample  %>% 
  get_ci(level = 0.95, type = "percentile")

infer_ci_95 

```

```{r}
infer_ci_95_visualised  <- infer_resample %>% 
  visualise() +
  shade_confidence_interval(endpoints = infer_ci_95)

infer_ci_95_visualised
```

# Using bootstrap distribution to calculate a 99% CI for mean(lot_area)

```{r}
infer_ci_99 <- infer_resample  %>% 
  get_ci(level = 0.99, type = "percentile")

infer_ci_99 
```

The CI is broader for the 99%. This makes sense because 

# 

```{r}
mean_lot_area_point_estimate <- infer_resample %>%
  summarise(mean(stat))
```

