---
title: "R Notebook"
output: html_notebook
---

# Loading in libraires 
```{r}
library(tidyverse)
library(janitor)
library(modelr)
library(broom)
library(ggfortify)
```

# Reading in data and plotting data 

```{r}
project <- read_csv("data/project_management.csv")  
  
```

```{r}
project_plot <- project %>% 
  ggplot(aes(x = estimated_length, y = actual_length)) +
  geom_point()+
  geom_text(aes(label = 1:nrow(project)))

project_plot
```

Looking at the plot it appears point labelled "5" is an outlier and has high leverage - potentially influencial
"28" and "18" look like outliers but less influential 

# Regressing actual_length on estimated_length and plotting cooks distance to  

```{r}
plot(lm(actual_length ~ estimated_length,data = project))
```

# Obtaining the intercept and regression coefficient of variable estimated_length for a simple linear model fitted to data omitting one non-influential outlier points
 
 Values when fitting the full data set
 
```{r}
model_1 <- lm(formula = actual_length ~ estimated_length, data = project)

model_1
```


Coefficients:
     (Intercept)  estimated_length  
           1.416             1.223 
           

When "18" point is ommitted.

```{r}
project_2 <- project %>%
  filter(X1 != 18) 
```

```{r}
model_2 <- lm(formula = actual_length ~ estimated_length, data = project_2)

model_2
```


Coefficients:
     (Intercept)  estimated_length  
           1.591             1.221  


The intercept and estimated_length values have changed after omitting "18" the non influnetial data point. intercept has increased and estimated_length has gone down.


#Plotting all the data points, the regression line omitting "18" and the regression line for the full data set

```{r}
project %>%
  ggplot(aes(x = estimated_length)) +
  geom_point(aes(y = actual_length)) +
  geom_abline(
    intercept = model_1$coefficients[1],
    slope = model_1$coefficients[2],
    col = "red"
  ) +
   geom_abline(
    intercept = model_2$coefficients[1],
    slope = model_2$coefficients[2],
    col = "blue"
  )
```

There appears to be a slight difference between the lines but the lines more or less run the same. 


# Obtaining the intercept and regression coefficient of variable estimated_length for a simple linear model fitted to data omitting influential outlier point

```{r}
project_3 <- project %>%
  filter(X1 != 5) 
```

```{r}
model_3 <- lm(formula = actual_length ~ estimated_length, data = project_3)

model_3
```

Coefficients:
     (Intercept)  estimated_length  
           4.381             1.011  

The intercept and estimated_length values have changed after omitting "5" the influnetial data point. intercept has increased and estimated_length has gone down by even greater values than the non-influential outliers


# Plotting all the data points, the regression lines omitting "18" and "5" and the regression line for the full data set

```{r}
project %>%
  ggplot(aes(x = estimated_length)) +
  geom_point(aes(y = actual_length)) +
  geom_abline(
    intercept = model_1$coefficients[1],
    slope = model_1$coefficients[2],
    col = "red"
  ) +
   geom_abline(
    intercept = model_2$coefficients[1],
    slope = model_2$coefficients[2],
    col = "blue"
  ) +
   geom_abline(
    intercept = model_3$coefficients[1],
    slope = model_3$coefficients[2],
    col = "green"
  )
```

# Returning to the fitted model for the complete data set (model_1) and commenting on the Residuals vs Fitted, Normal Q-Q and Scale-Location diagnostic plots

```{r}
autoplot(model_1)
```

Res vs Fit - residuals aren't independant and are not scattered around 0 - there's a pattern
normal qq - all residuals are close to the line
scale- location - line seems to be fairly horizontal, residuals are in fixed position above x axis

The res v fit suggests the regression assumptions are not reasonably satisfied
