---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(tidyverse)
library(gtools)
```

# Titanic Survivors Task

```{r}
titanic_tibble <- tibble(
  passenger = c(488, 812),
  crew = c(215, 703), 
)

titanic_tibble <- titanic_tibble %>%
    mutate(total_deaths = rowSums(.), 
           status = c("survived", "died")) %>%
  select(status, passenger, crew, total_deaths )

titanic_tibble 
```


Total people on board

```{r}
survived_titanic <- 488 + 215
died_titanic <- 812 + 703

total_people <- survived_titanic + died_titanic

total_people
```


Crew Probability
```{r}
crew_prob <- titanic_tibble %>% 
  summarise(crew_total = sum(crew)) %>% 
  mutate(probaility = crew_total / total_people) %>% 
  select(probaility)

crew_prob
```

Passenger Probability

```{r}
passenger_prob <- titanic_tibble %>% 
  summarise(pass_total = sum(passenger)) %>% 
  mutate(probaility = pass_total  / total_people)%>% 
  select(probaility)


passenger_prob
```

Crew and survived Probability

```{r}
crew_and_survived <- titanic_tibble %>%
  select(status, crew) %>%
  filter(status == "survived") %>%
  mutate(probaility = crew / total_people) %>%
  select(probaility)

crew_and_survived
```

Passenger and survived Probability

```{r}
pass_and_survived <- titanic_tibble %>%
  select(status, passenger) %>%
  filter(status == "survived") %>%
  mutate(probaility = passenger / total_people) %>%
  select(probaility)

pass_and_survived

```

Conditional - prob(A∣B)=prob(A∩B)/prob(B)

```{r}

probability_crew_survived <- crew_and_survived / crew_prob
probability_crew_survived
```

```{r}
probability_passenger_survived <- pass_and_survived / passenger_prob
probability_passenger_survived
```

Passengers were more likely to survive



# Probability Quiz

#1. If I roll a six-sided dice once, what is the probability of getting a number less than 5?

Probability is 0.67 or 2/3 - as there only numbers 1, 2, 3, 4 are less than 5 on a six-sided dice (4/6). Simplified 4/6 = 2/3 (0.67 rounded in decimal form).

#2. How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}

three_dice_sample_space <- rolldie(3) %>% 
  summarise(sample_space = nrow(rolldie(3)))

three_dice_sample_space

```

Sample space of 3 dice is 18

#3.  In an experiment where we roll four dice:


```{r}
s_4_dice <- rolldie(4) 
s_4_dice
  
```

# i - How many outcomes are in the event total_equals_5_6_7 (i.e. the total of the four dice equals 5, 6 or 7)?
```{r}

total_equals_5_6_7 <- s_4_dice %>%
  summarise(num_outcomes = rowSums(. == 5|6|7)) %>% 
  head(1)

total_equals_5_6_7
```

4 outcomes were the roll of 4 dice equals 5, 6, or 7

# ii - What is the probability of this event?

prob(event) = num(event outcomes) / num (all outcomes)

```{r}
s_4_dice_total_outcomes <-  nrow(rolldie(4))
s_4_dice_total_outcomes

prob_total_equals_5_6_7 <- 4 / 1296

prob_total_equals_5_6_7 

```

# iii - What assumptions do we apply in answering part ii?

The assumption is that all outcomes are equally likely 

# 4. Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

```{r}

group_heights <- c(100, 101, 102, 103, 104)

seats <- as_tibble(
  
  permutations(n = 5,
               r =5,
               v = group_heights,
               repeats.allowed = FALSE)
  
)

seats

ordered_seats <- seats %>% 
  filter(V1 == 104, V2== 103, V3 == 102, V4 == 101, V5 == 100 )

```

```{r}
prob_ordered_seats <- nrow(ordered_seats) / nrow(seats)
prob_ordered_seats
```

Probability is 0.008

# 5. I roll six dice:

# i What is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?

```{r}
six_dice_in_order <- rolldie(6) %>% 
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = TRUE))


prob_six_dice_in_order <- nrow(six_dice_in_order) / nrow(rolldie(6))

prob_six_dice_in_order

```

Probabilty is 2.143347e-05

# ii 

```{r}
six_dice_unordered <- rolldie(6) %>% 
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = FALSE))


prob_six_dice_unordered <- nrow(six_dice_unordered) / nrow(rolldie(6))

prob_six_dice_unordered 
```

Probabilty is 0.015