---
title: "R Notebook"
output: html_notebook
---

# Loading in libraries

```{r}
library(tidyverse)
library(ggthemes)
library(CodeClanData)
library(viridis)
library(wesanderson)
```

# 3.1 Comparison

## "late_deliveries" Plot

```{r}

view(late_deliveries)
dim(late_deliveries)
colnames(late_deliveries)
```

Depicts the number of late deliveries between 2015 and 2019. Line plot the best as there is only 2 variables and shows a how late deliveries change over time.

```{r}
ggplot(late_deliveries) +
  aes(x = date, y = late_deliveries) +
  geom_line(aes(colour = "red")) +
  labs(x = "\n Year",
       y = "\n Number of Late Deliveries",
       title = "Number of late deliveries between 2015 and 2019") +
  theme(panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue",
                                size = 0.5, linetype = "solid"),
  panel.grid.major = element_line(size = 0.5, linetype = 'solid',
                                colour = "white"), 
  panel.grid.minor = element_line(size = 0.25, linetype = 'solid',
                                colour = "white")) +
  scale_colour_hue(guide = FALSE)

  
```


## "recovery_times" Plot

```{r}
view(recovery_times)
dim(recovery_times)
colnames(recovery_times)

```

```{r}

prognosis_labelled <- recovery_times %>% 
  rename(Prognosis = prognosis) %>% 
  mutate(treatment_group = recode(treatment_group, 
                       "treatment_a" = "Treatment A", 
                       "treatment_b" = "Treatment B",
                       "control" = "Control"))

prognosis_labelled

ggplot(prognosis_labelled, aes(x=treatment_group, y=recovery)) +
    geom_bar(stat="identity", aes(fill=Prognosis)) +
  facet_wrap(~Prognosis) +
  coord_flip() +
  labs(x = "\n Treatment Group",
       y = "\n Recovery Period",
       title = "The recovery time for each treatment group split by prognosis") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank()) +
  scale_fill_manual(values = wesanderson::wes_palette("Rushmore1", n = 4))
  

```

## "fitness_levels" Plot

```{r}
view(fitness_levels)
dim(fitness_levels)
colnames(fitness_levels)
```


```{r}
ggplot(fitness_levels, aes(x=child, y = fitness_score)) +
    geom_bar(stat="identity", aes(fill=group)) +
    coord_flip()+
    labs(x = "\n Child",
       y = "\n Fitness Score",
       title = "Fitness score for each child and coloured by group") +
  theme(legend.background = element_rect(color = "steelblue", linetype = "solid")) +
  scale_colour_discrete(name="Group")
  

```


# 3.2 Connection
## blood_pressure

```{r}
colnames(blood_pressure)
view(blood_pressure)
```


```{r}
ggplot(blood_pressure, aes(systolic_blood_pressure_mm_hg, daily_saturated_fat_intake_mg)) +
  geom_point(shape=18, color="red") +
  labs(x = "\n Systolic Blood Pressure (mm,hg)",
       y = "\n Daily Saturated Fat Intake (mg)",
       title = "\n Systolic Blood Pressure (mm,hg) v Daily Saturated Fat Intake (mg)") +
  geom_smooth(method=lm,
             color="darkblue",
             se = FALSE) +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## car_use

```{r}
colnames(car_use)
view(car_use)

```

```{r}
percent_car_use <- car_use %>%
  mutate(car_use_percent = round((car_use_percent * 100), digits = 1)) %>% 
  mutate(air_so2_concentration_p = round((air_so2_concentration_ppm * 1000000), digits = 1))

percent_car_use
  
```

```{r}
ggplot(percent_car_use, aes(x = air_so2_concentration_ppm , y = car_use_percent)) +
  geom_point(aes(colour = city, size = population)) +
  scale_x_continuous(trans = "log10") +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "\n air CO2 concentration (ppm)",
       y = "\n Car Use Percentage (%)",
       title = "\n Air CO2 Concentration v Car Use Percentage for each City")
  

  
```

#3.3 Distribution

## "d20_outcomes"

```{r}
colnames(d20_outcomes)
view(d20_outcomes)
```

```{r}
ggplot(d20_outcomes) +
  aes(x = outcome) +
  geom_histogram(bins = 10, col = "white", fill = "dark green") + 
  theme_light() + 
  labs(y = "count", x = "outcome")
```

## "d20x5_outcomes"

```{r}
colnames(d20x5_outcomes)
view(d20x5_outcomes)
```

```{r}
ggplot(d20x5_outcomes) +
  aes(x = outcome) +
  geom_histogram(bins = 6, col = "white", fill = "dark green") + 
  theme_light() + 
  labs(y = "count", x = "outcome")
```

# 3.4 Composition

## "pension_surplus" plot

```{r}
colnames(pension_surplus)
view(pension_surplus)
```

## pension_liabilities plot

```{r}
colnames(pension_liabilities)
view(pension_liabilities)
```

```{r}
pension_surplus_reworked2
```

