---
title: "R Notebook"
output: html_notebook
---

#Loading the RPostgresql library 

```{r}

library(RPostgreSQL)

```

# Sourcing in my credentials file.

```{r}

home_location <- Sys.getenv("HOME")
credentials_location <- paste0(home_location,"/credentials.R")
source(credentials_location)

```

# Connecting the omni_employees database to R

```{r}

db_connect <- dbConnect(drv = PostgreSQL(),
                        user = postgresql_username,
                        password = postgresql_password,
                        dbname = "omni_employees",
                        host = "localhost")

```

#Removing username and password as connection has been established 

```{r}

rm(postgresql_username, 
    postgresql_password)

```


# Listing the table names within the database

```{r}

dbListTables(conn = db_connect)

```

#Using dbGetQuery to view the employees table 

```{r}
dbGetQuery(conn = db_connect,
           statement = "SELECT * FROM employees")
```

# Listing the column names within the employees database

```{r}

dbListFields(conn = db_connect,
             name = "employees")

```

# 1. Finding all the employees who work in the ‘Human Resources’ department.

```{sql connection=db_connect}

SELECT 
  *
FROM 
  employees
WHERE 
  department = 'Human Resources'

```

# 2. Finding the first name, last name, and country of the employees who work in the ‘Legal’ department.

```{sql connection=db_connect}

SELECT 
  first_name,
  last_name,
  country
FROM 
  employees
WHERE 
  department = 'Legal'

```

# 3.Counting the number of empployees based in Portugal. 

```{sql connection=db_connect}

SELECT 
COUNT
  (*) AS portuguese_employees_num
FROM 
  employees
WHERE 
  country = 'Portugal' 

```

# Viewing the pay_details table

```{sql connection=db_connect}

SELECT * FROM pay_details

```

# Finding the number of pay_detail records without a local account number

```{sql connection=db_connect}

SELECT 
COUNT
  (*) AS missing_local_account_num
FROM 
  pay_details
WHERE 
  local_account_no IS NULL 

```

# Finding pay_details records lacking both a local_account_no and iban number. No records have both missing.

```{sql connection=db_connect}

SELECT
  *
FROM 
  pay_details
WHERE 
  (local_account_no IS NULL) AND (iban IS NULL)

```

# List of employees first names and last names ordered alphabetically by their last name (put any nulls last).

```{sql connection=db_connect}

SELECT 
  last_name,
  CONCAT(first_name, ' ', last_name) AS names_alphabetically
FROM 
  employees
ORDER BY 
  last_name ASC NULLS LAST 

```

# List of employees first names, last names and countries, ordered alphabetically first by their country and second by their last name.

```{sql connection=db_connect}

SELECT 
  first_name
  last_name,
  country,
  CONCAT(country, ' - ', first_name, ' ', last_name) AS country_and_names_alphabetically
FROM 
  employees
ORDER BY 
  country ASC NULLS LAST,
  last_name ASC NULLS LAST
  
```

# List of the top 10 highest paid employees.

```{sql connection=db_connect}

SELECT 
  *
FROM 
  employees
ORDER BY 
  salary DESC NULLS LAST
LIMIT 
  10

```

# Finding the first name, last name and salary of the lowest paid of the employees in Hungary

```{sql connection=db_connect}

SELECT 
  first_name,
  last_name,
  salary
FROM
  employees
WHERE 
  country = 'Hungary'
  ORDER BY 
  salary ASC NULLS LAST

```

# Finding how many employees have first names beginning with F

```{sql connection=db_connect}

SELECT 
COUNT
  (*) AS num_of_f_employees
FROM 
  employees
WHERE 
  first_name LIKE 'F%'

```

#Finding full detials of employees who have a 'yahoo" email address

```{sql connection=db_connect}

SELECT 
  *
FROM 
  employees
WHERE 
  email LIKE '%yahoo%'

```

# Finding the number of pension enrolled employees not based in either France or Germany

```{sql connection=db_connect}

SELECT 
COUNT
  (*) AS num_employees_in_pscheme_exc_fra_ger
FROM 
  employees
WHERE 
  (pension_enrol = TRUE) 
  AND (country != 'France' AND country != 'Germany')

```

# Finding breakdown of the numbers of employees enrolled, not enrolled and whose enrollment status is unknown in the corporation pension scheme.

```{sql connection=db_connect}

SELECT
  pension_enrol,
COUNT
  (id) 
FROM 
  employees 
GROUP BY 
  pension_enrol 

```

# Finding the maximum salary among employees in the Engineering department who work 1.0 full-time equivalent hours

```{sql connection=db_connect}
SELECT 
  salary
FROM 
  employees
WHERE 
  department = 'Engineering' AND fte_hours = 1.0
ORDER BY 
  salary DESC NULLS LAST
LIMIT 
  1
```

# Finding count by department of the employees who started work with the corporation in 2003.

```{sql connection=db_connect}

SELECT
  department,
COUNT
  (id) AS num_2003_employees
FROM 
  employees 
WHERE 
  start_date 
    BETWEEN '2003-01-01' AND '2003-12-31'
GROUP BY 
  department 

```

# Disconnecting from omni_employees database

```{r}

dbDisconnect(db_connect)

```



















