---
title: "R Notebook"
output: html_notebook
---

# Loading in libraries

```{r}
library(tidyverse)
library(janitor)
library(ggiraphExtra)
library(GGally)
library(fastDummies)
```

# 1. Reading in data and exploring 

```{r}
diamonds <- read_csv("diamonds.csv")

glimpse(diamonds)
```

# 2. Using ggpairs() to identify if there are strong correlations between x, y, z variables and carat variable

```{r}
ggpairs(diamonds)
```

It appears there is a strong correlation between the letter variables and carat

# 3. Dropping x, y, z variables (as carat can be used to describe size)

```{r}
diamonds_clean <- diamonds %>% 
  select(-c("x", "y", "z"))
```

# 4. (i) Using ggpairs() to determine whether there is a storng correlation between price and predictor variables and then (ii) plotting any significant correlations found

```{r}
ggpairs(diamonds_clean)
```

Very strong correlation between price and carat (0.922)

```{r}
diamonds_clean %>% 
  ggplot() +
  aes(x = carat, y = price) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Much smaller correlation value between price and table (0.127)

```{r}
diamonds_clean %>% 
  ggplot() +
  aes(x = table, y = price) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

Negative correlation value for depth and price (-0.307) 

```{r}
diamonds_clean %>% 
  ggplot() +
  aes(x = depth, y = price) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

# Investigating the factor levels of the predictors - cut, clarity and color 

```{r}
diamonds_clean %>% 
  distinct(cut)

# its likely 5 dummy variables would be created for cut

diamonds_clean %>% 
  distinct(clarity)

# its likely 8 dummy variables would be created for clarity

diamonds_clean %>% 
  distinct(color)

# its likely 7 dummy variables would be created for color
```

# Using fastDummies and dummy_cols() to generate dummies for the predictors - cut, clarity and color 

```{r}
diamonds_dummy <- diamonds_clean %>% 
  dummy_cols(select_columns = c("cut", "clarity", "color"),
             remove_first_dummy = TRUE,
             remove_selected_columns = TRUE)

diamonds_dummy
```

# (i) Simple linear regression - regressing price on carat and then checking the regression diagnostics 

```{r}


model_diamonds <- lm(price ~ carat, data = diamonds_clean)

summary(model_diamonds)
```

p-value is very low 0.00000000000000022 

RSE is 1549 on 53938 degrees of freedom 

# Running a regression with one or both of the predictor and response variables log() transformed and recheck the diagnostic

```{r}
model_diamonds2 <- lm(log(price) ~ log(carat) + log(depth) + log(table), data = diamonds_clean)

summary(model_diamonds2)
```

RSE has gone down to 0.2584 on 53936 degrees of freedom from 1549 on 53938 degrees of freedom

Multiple R-squared up from 0.8493 to 0.9351 

p-value is lower 2.2e-16

There is an imporvement so adding the log(depth) and log(table) is useful 


# (iii) Using log() transformations of both predictor and response. Then adding a single categorical predictor into the model to determine which categorical predictor is best.

## cut 
```{r}
model_diamonds3 <- lm(log(price) ~ log(carat) + cut, data = diamonds_clean)

summary(model_diamonds3)
```

Multiple R-squared:  0.9371

## clarity
```{r}
model_diamonds4 <- lm(log(price) ~ log(carat) + clarity, data = diamonds_clean)

summary(model_diamonds4)
```

Multiple R-squared:  0.9654

## color
```{r}
model_diamonds5 <- lm(log(price) ~ log(carat) + color, data = diamonds_clean)

summary(model_diamonds5)
```

Multiple R-squared:  0.9454


Clairty is the best categorical predictor as it has the highest R-squared value 


# Interpreting the fitted coefficients for the levels of clarity - categorical predictor. Which level is the reference level? Which level shows the greatest difference in price from the reference level?

```{r}
diamonds_clean %>% 
  ggplot(aes(x = log(price))) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 4)
```

```{r}
model_diamonds6 <- lm(log(price) ~ clarity, data = diamonds_clean)

summary(model_diamonds6)
```

```{r}
diamonds_clean %>% 
  ggplot(aes(y = log(price), x = clarity)) +
  geom_boxplot()
```

The reference level is I1 (the first level in graph above) and the greatest difference in price from the reference level is IF - this is because from the boxplots above the two median values for these categories appear to be furthest away from each other and price has been logged)
  
   
  
  
